---
title: "Data analysis Pilot 1"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: yes
    latex_engine: lualatex
    dev: cairo_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(markdown)
library(tidyverse)
library(lme4)
library(gridExtra)
library(afex)
library(cowplot)
```

## Fit algorithm definition
```{r}

fit_algo <- function(new_call,
                     newData = NULL,
                     type = "lmer")
{
  if (type != "lmer" && type != "glmer")
  {
    stop("type must be lmer or glmer")
  }
  new_call["start"] <- NULL
  new_call["control"] <-
    parse(text = paste0(
      type,
      'Control(optimizer ="bobyqa", optCtrl = list(maxfun=1000000))'
    ))
  error_msg <- tryCatch({
    rval <- eval(new_call)
    if (length(rval@optinfo$conv$lme4) != 0)
    {
      "convergence"
    }
    else
    {
      "ok"
    }
  }, error = function(e) {
    "error"
  })
  count <- 0
  while (error_msg != "ok" && count < 4)
  {
    if (error_msg == "convergence")
    {
      new_call["start"] <- parse(text = "list(fixef = fixef(rval))")
    }
    count <- count + 1
    error_msg <- tryCatch({
      rval <- eval(new_call)
      if (length(rval@optinfo$conv$lme4) != 0)
      {
        "convergence"
      }
      else
      {
        "ok"
      }
    }, error = function(e) {
      "error"
    })
  }
  new_call["start"] <- NULL
  new_call["control"] <-
    parse(text = paste0(
      type,
      'Control(optimizer ="Nelder_Mead", optCtrl = list(maxfun=1000000))'
    ))
  count <- 0
  while (error_msg != "ok" && count < 5)
  {
    count <- count + 1
    error_msg <- tryCatch({
      rval <- eval(new_call)
      if (length(rval@optinfo$conv$lme4) != 0)
      {
        "convergence"
      }
      else
      {
        "ok"
      }
    }, error = function(e) {
      "error"
    })
    if (error_msg == "convergence")
    {
      new_call["start"] <- parse(text = "list(fixef = fixef(rval))")
    }
    
  }
  new_call["start"] <- NULL
  new_call["control"] <- NULL
  count <- 0
  while (error_msg != "ok" && count < 5)
  {
    count <- count + 1
    error_msg <- tryCatch({
      rval <- eval(new_call)
      if (length(rval@optinfo$conv$lme4) != 0)
      {
        "convergence"
      }
      else
      {
        "ok"
      }
    }, error = function(e) {
      "error"
    })
    if (error_msg == "convergence")
    {
      new_call["start"] <- parse(text = "list(fixef = fixef(rval))")
    }
    
  }
  if (!exists("rval"))
  {
    eval(new_call)
  }
  return(rval)
}

```

## Dataset loading

```{r}
datasets <-
  list.files(file.path("final_data"),
             pattern = "*.csv",
             full.names = TRUE)
datasets       <- lapply(datasets, read_csv, show_col_types = FALSE)

d <- datasets[[1]]
for (i in seq(from = 2, to = length(datasets)))
{
  d <- rbind(d, datasets[[i]])
}
```

## Dataset preparation

```{r}
d <- d %>% mutate(Subject = rep(seq_len(length(datasets)), each = 100))
```

Then we remove fixation times <=1500ms, scale the Agreement feature and take its square, and log-transform Fixation.Time

```{r}
d_final <- d %>% filter(Fixation.Time > 1500) %>%
  mutate(
    Fixation.Time.Log = log(Fixation.Time),
    Agreement.Scaled = (Agreement - mean(Agreement)) / sd(Agreement),
    Squared.Agreement.Scaled = Agreement.Scaled ^ 2,
    Index = as.factor(Index),
    Subject = as.factor(Subject)
  )
```

## Variables plot (violin & boxplot + density)

```{r}

violin_fixation <-
  ggplot(d_final, aes(x = "Fixation time", y = Fixation.Time)) +
  ylab("Fixation Time") + theme_minimal() +
  theme(legend.position = "none") +
  geom_violin(width = 1.2) + geom_boxplot(width = 0.1, alpha = 0.2)


violin_agreement <-
  ggplot(d_final, aes(x = "Agreement", y = Agreement.Scaled)) +
  ylab("Agreement (scaled)") + theme_minimal() +
  theme(legend.position = "none") +
  geom_violin(width = 1.2) + geom_boxplot(width = 0.1, alpha = 0.2)

density_fixation <-
  ggplot(d_final, aes(x = Fixation.Time)) + geom_density()

density_agreement <-
  ggplot(d_final, aes(x = Agreement.Scaled)) + geom_density()

grid.arrange(violin_fixation,
             violin_agreement,
             density_fixation,
             density_agreement,
             nrow = 2)

```

## Statistical models (log-tranformed, GLMM with Gamma-distributed errors and LMM with normal-distributed errors)

```{r}
log_call <- parse(
  text = 'lmer(Fixation.Time.Log ~ Agreement.Scaled +
                    Squared.Agreement.Scaled +
                    (1 | Index) +
                    (Agreement.Scaled + Squared.Agreement.Scaled | Subject),
                  data=d_final)'
)[[1]]

fit_log <- fit_algo(log_call)

summary(fit_log)

fit_call <- parse(text = 'lmer(Fixation.Time ~ Agreement.Scaled +
                    Squared.Agreement.Scaled +
                    (1 | Index) +
                    (1 | Subject),
                  data=d_final)')[[1]]

fit <- fit_algo(fit_call)

summary(fit)
```

### Fixed effects plots

```{r}
d_final$log_prediction <- exp(predict(fit_log, re.form = NA))
d_final$prediction <- predict(fit, re.form = NA)

colors <-
  c(
    "Log transformed" = "lightpink2",
    "Normal-assumed errors" = "lightgreen"
  )

ggplot(d_final) +
  geom_line(aes(x = Agreement.Scaled, log_prediction, color = "Log transformed"),
            linewidth = 1.5) +
  geom_line(aes(x = Agreement.Scaled, prediction, color = "Normal-assumed errors"),
            linewidth = 1.5) +
  theme_minimal_grid() +
  theme(text = element_text(size = 16)) +
  labs(y = "Fixation time", x = "Agreement", color = "Model") +
  scale_color_manual(values = colors)

plot <- ggplot(d_final) +
  geom_line(
    aes(x = Agreement.Scaled, prediction),
    color = "lightgreen",
    linewidth = 1.5
  ) +
  theme_minimal_grid() +
  theme(text = element_text(size = 16)) +
  labs(y = "Fixation time", x = "Agreement")
plot
ggsave(plot, filename = "study1_plot.png", type = "cairo")
```

```{r}
vertex <-
  (-summary(fit)[["coefficients"]][2]) / (2 * summary(fit)[["coefficients"]][3])

pred <-
  predict(
    fit,
    re.form = NA,
    newdata = tibble(
      Agreement.Scaled = c(
        min(d_final$Agreement.Scaled),
        vertex,
        max(d_final$Agreement.Scaled)
      ),
      Squared.Agreement.Scaled = Agreement.Scaled ** 2
    )
  )

table <- round(matrix(
  c(
    vertex,
    min(d_final$Agreement.Scaled),
    max(d_final$Agreement.Scaled),
    pred[2] - pred[1],
    pred[2] - pred[3]
  ),
  ncol = 1,
  byrow = TRUE
), 3)

rownames(table) <-
  c(
    'Vertex position',
    'min agreement',
    'max agreement',
    'diff min vs. vertex',
    'diff max vs. vertex'
  )
colnames(table) <- c('Results')
table <- as.table(table)
table
```